\speculake: Hiding Malware in Speculative Execution

### 2018-05-07 paper/implementation.tex
VM implementation updates

\subsection{Virtual Machine}

\subsubsection{SPASM}
\label{subsubsec:spasm}
Constructing an emulator making use of the speculative primitive requires  
a trade off in expresive capability versus speed. 
We have implemented SPASM as a model using two pseudo-registers, and six 
bit instruction length which allows for a relatively direct programming model 
in which structured values can be entered into memory locations before making 
a systemcall.

To acheive a balance with speed the number of bits in each instruction is both 
fixed and minimized.  A variable length instruction would require that the 
\texttt{Flush+Reload} stage search the maximum number of bits on each round, and each 
aditional bit doubles the search space that the \texttt{Flush+Reload} stage must 
traverse.  So every bit shorted effectively doubles the throughput of the 
emulator, and there is effectively no advantage to allowing variable length 
instructions.

Using this model we have implemented multiple example programs that can be run
as encrypted payloads in an \speculake payload. A \textit{HelloWorld} program 
that prints to stdout. A \textit{FizzBuzz} program that demonstrates control 
flow and arithmetic operations while printing to stdout. And finally,
a \textit{ReverseShell} program that opens and connects to a socket before  
duplicating I/O and executing a local shell.

Expected performance for a given SPASM binary will vary given multiple factors, 
the most important of which is the \texttt{Flush+Reload} redundancy. This can be tuned
based on the specific trigger, as it is used to establish confidence that a signal 
has been identified in the indicess returned from the "speculative world". 
Thus accomplishing a task using SPASM generally equates to:

\begin{lstlisting}
    Redundancy * Probe_Space * Num_Instr
\end{lstlisting}

For example, the \textit{ReverseShell} program makes six system calls to open a socket,
connect to it, duplicate I/O, and execve a specified program. This requires 355
SPASM instructions to complete. The probe space for the 6-bit SPASM ISA requires 
a traversal of an array od size 64. When using a redundancy of 256 to establish
confidence in the signal the entire \textit{ReverseShell} program takes under 7 
seconds to run once triggered.

@@

\subsubsection{ARM}
\label{subsubsec:arm}
While the custom emulator that we developed gives a much higher level abstraction to 
a developer, it still requires programs to be written in custom assembly 
language. To demonstrate that the extensibility of this model we also implemeted a wrapper
around an ARM emulator, such that programs can be written in a high level languages 
such as C or C++, statically compiled to binaries for the ARM emulator, and deployed 
as an encrypted data sections to be decrypted at run-time.

% ARM Wrapper
%   - memory
%   - single instruction vs block

We chose to implement a wrapper around ARM as it has a 32-bit fixed instruction
length. However this causes a significant hit to the speed at which operations 
can be performed as the probe space is increased by a factor of four in 
comparison with SPASM,  and we must complete 4 of these transactions to 
retreive a single ARM opcode. Trading more expressive operations for speed
results in a 16x slowdown when considering per-instruction throughput.

\begin{lstlisting}
    4 * Redundancy * Probe_Space * Num_Instr
\end{lstlisting}

>>>>

\subsection{Virtual Machine}
\label{subsec:spasm}

% \subsubsection{SPASM}
% \label{subsubsec:spasm}
Constructing an emulator making use of the speculative primitive requires  
a trade off in expresive capability versus speed. 
We have implemented SPASM as a model using two pseudo-registers, and six 
bit instruction length which allows for a relatively direct programming model 
in which structured values can be entered into memory locations before making 
a systemcall.

To acheive a balance with speed the number of bits in each instruction is both 
fixed and minimized.  A variable length instruction would require that the 
\texttt{Flush+Reload} stage search the maximum number of bits on each round, and each 
aditional bit doubles the search space that the \texttt{Flush+Reload} stage must 
traverse.  So every bit shorter effectively doubles the throughput of the 
emulator, and there is effectively no advantage to allowing variable length 
instructions. Details of the SPASM instruction set can be found in 
Appendix~\ref{appendix:spasm}.

Using this model we have implemented multiple example programs that can be run
as encrypted binaries in an \speculake payload. A \textit{HelloWorld} program 
that prints to stdout. A \textit{FizzBuzz} program that demonstrates control 
flow and arithmetic operations while printing to stdout. And finally,
a \textit{ReverseShell} program that opens and connects to a socket before  
duplicating I/O file descriptors and executing a local shell.

Expected performance for a given SPASM binary will vary given multiple factors, 
the most important of which is the \texttt{Flush+Reload} redundancy. This can be tuned
based on the specific trigger, as it is used to establish confidence that a signal 
has been identified in the indicess returned from the "speculative world". 
%Thus accomplishing a task using SPASM generally equates to:

% \begin{lstlisting}
%     Redundancy * Probe_Space * Num_Instr
% \end{lstlisting}

For example, the \textit{ReverseShell} program makes six system calls to open a socket,
connect to it, duplicate I/O file descriptors, and perform an execve
systemcall to open a shell. This requires 355 SPASM instructions to complete. 
The probe space for the 6-bit SPASM ISA requires a traversal of an array of size 64.
When using a redundancy of 256 to establish confidence in the signal the entire 
\textit{ReverseShell} program takes under 7 seconds to run once triggered.

% TODO - REFERENCE APPENDIX for ISA

@@

### 2018-05-07 paper/payloads.tex
VM implementation updates

To accommodate both developers as well as the unique constraints 
enforced by the speculative primitive, a custom emulation solution is required. 
we have approached this challenge by creating a custom instruction set -- 
SPASM -- which accommodates the speculative primitive specifically. This 
provides a higher level of abstraction around the speculative primitive
that we have used to implement various payloads.

@@

% REFERENCE FIGURES
% - Constraints enforced by speculative primitive
%       - more bits = slower
%       - less bits = less expressive
%       - must be computable within speculative ROB limit (~150 instrs) (decrypt only)
%   
% - optimizations for weird-machine computations

% --- VM
%   - Wrappers around traditional ISA VMs
%   - SPASM  custom emulator made f
%%%%%%%%%%%%%%

>>>>

To accommodate both developers as well as the unique constraints 
enforced by the speculative primitive, a custom emulation solution is required. 
We have created a custom emulator and instruction set -- SPASM -- which 
accommodate the speculative primitive specifically. This 
provides a higher level of abstraction around the speculative primitive
that we have used to implement various executable binaries.

@@

While the custom emulator that we developed gives higher level abstraction to
an author, it still requires programs to be written in a custom assembly language. 
We note that the \speculake model is not intrinsically linked to the SPASM emulator. 
A wrapper could be implemented around other existing emulators to incrementally 
construct instructions 6 or 8 bits at a time (depending on the channel width). 
This would allow for encrypted payloads to be written in higher level languages
like C or C++, although there would be a significant performance reduction to 
using this type of wrapper.
